Sketch Recognition using Particle Swarm Optimizations

### 2009-05-27 trunk/paper2/Paper2.tex
fininshign the introduction

%\documentclass{article}
\documentclass[preprint,authoryear,10pt,3p]{elsarticle}
%\usepackage{spconf}
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage[dvips,pdftex]{graphicx}
\usepackage{subfigure}
%\usepackage{named}
%\usepackage[dvips,pdftex]{hyperref}

%\title{Sketch Recognition using Particle Swarm Optimizations}
%\twoauthors
%  {Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
%	{Signals Processing Group \\ Computers and Systems Department \\
%  Electronic Research Institute\\Cairo, Egypt\\
%	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,\\ samia@eri.sci.eg}
%  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
%	{Computer Engineering Department\\
%	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
%	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Sketch Recognition using Particle Swarm Optimizations}

@@

\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
\author[ERI]{Mahmoud Fakhr El Din}
\author[ERI]{Samia Mashali}
\author[FAC]{Magda Fayek}
\author[FAC]{Nevin Darwish}
\address[ERI]{Signals Processing Group, Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
\address[FAC]{Computer Engineering Department\\
	Faculty of Engineering,Cairo University\\Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{abstract}
%%This study 
Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).  % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

@@

%\maketitle
%\begin{abstract}
%%This study 
%Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).  % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.  
%\end{abstract}

@@

Computer research nowadays aims at facilitating computer human interaction in every way. Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces. However, Using sketches as input to a computer system requires the system to understand the drawn sketches. In a truly freely sketch system this is not a trivial problem due to the sloppy and messy way users draw sketches. Sketches here stand for any hand drawing the user draws using digital stylus or similar device. Normally, sketches are composed of a set of strokes. The time between pen-up and pen-down events identify each stroke. The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}. The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods. Most stroke based systems either represent symbols as gestures or as a  hierarchical shape. Gestures are pen strokes drawn by the user that can immediately recognized by the system. A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}. Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them. Systems that employ such representation must first pre-process strokes to break them into geometric primitives. Recognition can then be treated as sub graph matching problem.

 In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties (i.e. ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}. Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes. In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw.

 Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}. This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties. \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

 Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols. Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.

In this paper we choose to implement a hybrid method between shape property and stroke based approaches. Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use. Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.

@@

%This paper introduces a new sketch recognition system, which uses  % uses the computed feature vector to classify the segments into one of the previously trained classes.

 The remaining of the paper is as follows; the literature review on stroke segmentation is in Section \ref{sec:review}. Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization. Section \ref{Sysdisc} describes the system in details.  Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively. Section \ref{sec:Experiments} describes the experiments that were preformed. We conclude in Section \ref{ConclusionandFutureWork}.

@@

\subsection{Preprocessing}
\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}. It was noted that time difference provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals as long as the pointing device is moving. Therefore, there are no new detected points when the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}). Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
 \begin{figure}[h]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}
 \begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}
Dominant points are characterized by lower speed values, higher curvature and direction values. However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information. Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.

>>>>

\documentclass[preprint,10pt,5p,twocolumn]{elsarticle}
%\documentclass[preprint,authoryear,10pt,5p,twocolumn]{elsarticle}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfigure}

%% for the old format  ( remove this and the abstract and the bibgraphy style. )
%\documentclass{article}
%\usepackage{spconf}
%\usepackage{amsmath}
%\usepackage[dvips,pdftex]{graphicx}
%\usepackage{named}
%\usepackage[dvips,pdftex]{hyperref}

%\title{Sketch Recognition using Particle Swarm Optimizations}
%\twoauthors
%  {Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din\sthanks{Associate Professor}, Samia Mashali\sthanks{Professor}}
%	{Signals Processing Group \\ Computers and Systems Department \\
%  Electronic Research Institute\\Cairo, Egypt\\
%	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,\\ samia@eri.sci.eg}
%  {Magda Fayek\sthanks{Associate Professor}, Nevin Darwish\sthanks{Professor}}
%	{Computer Engineering Department\\
%	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
%	magdafayek@gmail.com,ndarwish@ieee.org}
%\maketitle

\begin{document}
\begin{frontmatter}
\title{Sketch Recognition using Particle Swarm Optimizations}

@@

\author[ERI]{Maha El Meseery\footnote{Crossponding author}}
\author[ERI]{Mahmoud Fakhr El Din}
\author[ERI]{Samia Mashali}
\author[FAC]{Magda Fayek}
\author[FAC]{Nevin Darwish}
\address[ERI]{Signals Processing Group, Computers and Systems Department, 
  Electronic Research Institute, Cairo, Egypt\\melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg, samia@eri.sci.eg}
\address[FAC]{Computer Engineering Department,
	Faculty of Engineering,Cairo University, Cairo, Egypt\\
	magdafayek@gmail.com,ndarwish@ieee.org}
\begin{abstract}
%%This study 
Sketches and drawings are widely used as simple methods of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to form symbols which are further identified using a Support Vector Machines (SVM) classifier. Two versions of PSO segmentation algorithms were implemented. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that both PSO algorithms produce better segmentation than most current systems. This segmentation improvement contributed in getting more accurate recognition results on all the datasets tested. The results prove that the system is effective specially in datasets with low diversity in shapes, i.e., shapes are very simillar, as in Logic Design dataset (LD-DS). %The results prove the effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS).% effectiveness of the proposed system on various dataset, especially datasets with low shape diversity as in Logic Design dataset (LD-DS). %Result show that the system is especially effictive on dataset with % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier which further improves the final recognition accuracy.

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Symbol Recognition, Particle Swarm Optimization, Sketch Recognition, Geometerical Features, Polygon approximation
%% PACS codes here, in the form: \PACS code \sep code

@@

%\begin{abstract}
%%This study 
%Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DB), Electrical Design (EL-DB) and Logic Design (LD-DB) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).  % sketch are divided  Users draw symbols and sketchs using a set of pen strokes. 
%%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple %strokes. Usually, users draw strokes using pens. The sketch recognition system immediately interprets their %strokes as objects that can be easily manipulated. This paper uses Particle Swarm Optimization (PSO) to divide %the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to %formulate symbols which are further identified. The results show that using PSO improves segmentation results %which guide the symbol recognition phase. As for recognition we use Support Vector Machines (SVM) classifier %which further improves the final recognition accuracy.  
%\end{abstract}

@@

Computer research nowadays aims at facilitating computer human interaction in every way. Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces. However, using sketches as input to a computer system requires the system to understand the drawn sketches. In a truly freely sketch system this is not a trivial problem due to the sloppy and messy  way in which users draw sketches. Sketches here stand for any hand drawing the user draws using digital stylus or similar device. Normally, sketches are composed of a set of strokes. The time between pen-up and pen-down events identifies each stroke. The strokes are mainly the path of points, which the user draws between the pen-down and pen-up events while using the stylus or similar device.

Sketch recognition can be applied either using  Pattern Recognition (PR) based method or Artifical Intelligance (AI) based method. In AI based methods, Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them. Systems that employ such representation must first pre-process. strokes to break them into geometric primitives. Recognition can then is treated as an AI problem which can be solved using methods like sub graph matching, bayes belief network, and constrain statisfaction. A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,SketchRead2007}.  %Ladder30,AlvaradoFreedom42

Recognition based on AI methods as applied in \cite{SketchRead2007} has the principle problem of computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to be segmented and grouped to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols. Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch. This is because they assume that each stroke plays a role in the structure of the shape. 
%The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods.  
%Gestures are pen strokes drawn by the user that can immediately recognized by the system.
%role of which each stroke plays in the structure of the sketched symbol is used to determine shapes 
%. discribing the role  how user draws individal storkes  properties that discribe how users draw individual strokes in drawing the symbol.  to . The identification of individual stroke is used to determine  The symbols are recognized by identifing the individual strokes.
In PR based methods, systems compute properties that discribe symbols then use a classification method to recognize symbols. The properties used can either be based on gesture properties, global shape properties or image based properties. Gesture properties are set of features that represent how the user draws a symbol not how it looks like (e.g., identify a regtangle when a user draws a L corner). In gesture properties approcah, systems identify symbols by using gesutre properties of some of its  underlying strokes. System that uses gestures properties assume that users draw symbols in specific order and style \cite{gestureexample12,aideddesgin22}. In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties (e.g., ratio of the area of convex hull to the area of shape) relaxes the assumption and the role of each stroke in the symbol but it does not truly represent the appearance or fine details of the symbol \cite{DiagramOfflineConvexHull,Cali63}. Unlike shape properties based methods, image based methods disregard individual strokes and focus on the overall appearance of shapes. These system convertes the shapes into image and extracte image based features (e.g. local gradient feature)\cite{Oltmans07,imagetrainable48}. In \cite{imagetrainable48} used four different distance measures to to match user symbol with the prototype database. The prototype of a shape is represented as an image of the symbol that is normalized and scaled.  However, the problem is that the shape prototype must represent all transformation and variation the input shape can have.   %Image based methods are used in \cite{Oltmans07,imagetrainable48} where they computed a code book of 1000 features to recognize symbols.

Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}. This assumption allows the system to temporally group strokes and then recognizes them based on their strokes or global shape properties. Alvarado \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

In this paper, we choose to implement a hybrid method between shape property, guesture and hierarchical (AI method) based approaches. Our method attempts to avoid many of the problems faced by approaches based on individual guestures or global shape properties while trying to gain the additional information that the stroke based system uses. Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.

 %In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw. 
%Sketch recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}. The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods. Most stroke based systems either represent symbols as gestures or as a  hierarchical shape. Gestures are pen strokes drawn by the user that can immediately recognized by the system. A Hierarchical shape descriptions was widely used in sketch systems \cite{HierarchicalParsing7,Ladder30,AlvaradoFreedom42,SketchRead2007}. Symbols are represented using hierarchical levels of simpler geometric shapes or primitives and spatial relations between them. Systems that employ such representation must first pre-process strokes to break them into geometric primitives. Recognition can hen be treated as sub graph matching problem.

 %In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties (i.e. ratio of area of convex hull to area of shape) relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol\cite{DiagramOfflineConvexHull}. Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes. In general, the  appearance based methods use image processing techniques to recognize sketches that the users draw.

 %Methods based on gesture and global shape properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{Cali63,geometrydomain49}. This let them to temporally group strokes and then recognizes them based on their strokes or global shape properties. \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

 %Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols. Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.

%In this paper we choose to implement a hybrid method between shape property and stroke based approaches. Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use. Firstly, the system uses particle swarm optimization (PSO) to optimally segment user's strokes. Users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different PSO algorithms. The first segmentation algorithm is based on dividing the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm enhances its performance by adding curvature and other local information while segmenting strokes to achieve better results. To handle complex strokes, our system uses a second segmentation PSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a feature vector composed of different statistical, geometrical and composite features is deduced. The final classification step uses a SVM classifier to classify symbols into one of the previously trained classes based on the features vector computed.

@@

%This paper introduces a new sketch recognition system, which uses  % uses the computed feature vector to classify the segments into one of the previously trained classes.
%Sections \ref{Prepross}, \ref{seg} and \ref{sec:Recognition} provide details on preprocessing, segmentation and recognition process respectively.
 The remaining of the paper is as follows; the literature review on stroke segmentation is presented in Section \ref{sec:review}. Section \ref{sec:ParticleSwarmAlgorithm} introduces Particle Swarm Optimization. Section \ref{Sysdisc} describes the proposed system in details. Section \ref{sec:Experiments} describes the experiments that were preformed. Finally, the conclusion and future work is presented in Section \ref{ConclusionandFutureWork}.

@@

\subsection{Preprocessing}
\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}. It was noted that time difference provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals as long as the pointing device is moving. Therefore, there are no new detected points when the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}). Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}

 \begin{figure}[h]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}
 \begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}
Dominant points are characterized by lower speed values, higher curvature and direction values. However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information. Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.
